---
title: 'Reproducible Research: Peer Assessment 1'
author: "Álvaro Sánchez Rodríguez"
date: "September 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

#Introduction 

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Libraries 
```{r libraries, echo=FALSE}
libraries_load=c("ggplot2","dplyr","lubridate")
```

Before to begin with the assessment, we are loading libraries requiered. In this case we load: `r libraries_load`.

```{r, results='hide',message=FALSE}
ipak <- function(pkg){
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        if (length(new.pkg)) 
                install.packages(new.pkg, dependencies = TRUE)
        sapply(pkg, require, character.only = TRUE)
}

ipak(libraries_load)
```

## System Information
In order to replicate or reproduce the analyses, I think that it must be useful to supply system information:
```{r, echo=FALSE}
info=R.Version()
```
1. Platform: `r info$platform`
2. R version: `r info$version.string`
3. Date: `r Sys.Date()`

And as no english native speaker I decided to change my configuration to  by means of `en_US.UTF-8`:


```{r}
Sys.setlocale("LC_TIME", "en_US.UTF-8")
```


# Loading and preprocessing the data

First of all we established our working directory by means of `getwd()` and unzip and load data:
```{r, echo=FALSE}
setwd("/media/DATA2/cursos/R/Reproducible Research/Peer Assessment 1/RepData_PeerAssessment1/")
```
```{r}
unzip("activity.zip")
data_act=read.csv("activity.csv")
```

Then we can format and transform variables as we want. In this case I decided to transform dates to julian days.

```{r}
data_act$date=as.Date(data_act$date)
data_act$julian=julian(data_act$date,origin=min(data_act$date))
```

Now we can start to answer questions.

#Questions

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
by_day=group_by(data_act,julian)
steps_per_day=summarise(by_day,sumatory=sum(steps, na.rm=TRUE),average=mean(steps))
steps_per_day=filter(steps_per_day, sumatory!=0)
```

2. Make a histogram of the total number of steps taken each day
```{r, fig.align='center',fig.width=5, message=FALSE}
ggplot(data=steps_per_day,aes(x=sumatory))+
        geom_histogram(fill="cadetblue",colour="black")+
        ylab("Count")+xlab("Total number of steps per day")+
        theme_bw()+
        theme(panel.grid.minor=element_blank(),
              axis.text.x = element_text(colour="grey20",size=11,angle=0,hjust=.5,vjust=.5,face="bold"),
              axis.text.y = element_text(colour="grey20",size=11,angle=0,hjust=.5,vjust=.5,face="bold"),
              axis.title.y = element_text(colour="grey20",size=15,angle=90,hjust=.5,vjust=1,face="bold"),
              axis.title.x = element_text(colour="grey20",size=15,angle=0,hjust=.5,vjust=-0.2,face="bold"),
              legend.key=element_blank()
        )
```


3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(steps_per_day$sumatory)
median(steps_per_day$sumatory)
```

The mean and median of the total number of steps taken per day are **`r format(mean(steps_per_day$sumatory),scientific=F)`** and **`r median(steps_per_day$sumatory)`** respectively. 

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r, fig.align='center',fig.height=5}
by_interval=group_by(data_act,interval)
steps_per_interval=summarise(by_interval,media=mean(steps,na.rm=TRUE))

ggplot(data=steps_per_interval,aes(x=interval,y=media))+
        geom_line(colour="cadetblue")+
        ylab("Mean of steps taken every day")+xlab("Interval of the day (min)")+
        theme_bw()+
        theme(panel.grid.minor=element_blank(),
              axis.text.x = element_text(colour="grey20",size=11,angle=0,hjust=.5,vjust=.5,face="bold"),
              axis.text.y = element_text(colour="grey20",size=11,angle=0,hjust=.5,vjust=.5,face="bold"),
              axis.title.y = element_text(colour="grey20",size=15,angle=90,hjust=.5,vjust=1,face="bold"),
              axis.title.x = element_text(colour="grey20",size=15,angle=0,hjust=.5,vjust=-0.2,face="bold"),
              legend.key=element_blank()
        )
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maximum=filter(steps_per_interval,media==max(media))
maximum_period=seconds_to_period(maximum$interval*60)
maximum
```

The 5-minute interval which shows the maximum number of steps (on average across all days) is `r maximum$interval` which corresponds to `r paste(maximum_period)`.


## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as `NA`). The presence of missing days may introduce bias into some calculations or summaries of the data.



## Are there differences in activity patterns between weekdays and weekends?
